---
title: "Mimic CC RnaSeq"
author: Chase Mateusiak, chasem@wustl.edu
date: "`r Sys.Date()`"
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), "../docs", 'index.html')) })
output: 
  html_document
---

<<<<<<< HEAD
# plot the background and foreground for GSEA analysis
https://www.huber.embl.de/users/klaus/Teaching/DESeq2Predoc2014.html#gene-ontology-enrichment-analysis

# goStats has some analysis suggestions
https://bioconductor.org/packages/release/bioc/vignettes/GOstats/inst/doc/GOvis.pdf

# profilr -- combine chip and rnaseq
#https://bioconductor.org/packages/devel/bioc/vignettes/profileplyr/inst/doc/profileplyr.html

=======
>>>>>>> 8fa902069a054d712991ad24f35a0a2007b11ed5
# see this data set + chp on diaxic shift in gcr1 mutants
# https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE153842

# paper on lysine pathway
# https://pubmed.ncbi.nlm.nih.gov/16943623/

```{r notebook_setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  echo = TRUE,
  include = TRUE,
  cache = TRUE
)
```

# Setup environment

```{r setup, include=FALSE, echo=FALSE}
# note, kept here for convenience. Repeated so that it shows up in notebook
library(DESeq2)
library(gprofiler2)
library(patchwork)
library(tidyverse)
library(DT)
library(pheatmap)
library(here)
library(gridExtra)
library(grid)
```

```{r, include=FALSE, echo=TRUE}
library(DESeq2)
library(gprofiler2)
library(patchwork)
library(tidyverse)
library(DT)
library(pheatmap)
library(here)
library(gridExtra)

library(ggVennDiagram)
# note that to install this i needed
# sudo apt install libpq5=12.9-0ubuntu0.20.04.1
# sudo apt install libpq-dev=12.9-0ubuntu0.20.04.1
# sudo apt install libgdal-dev 
<<<<<<< HEAD

source(here("R/create_long_results_df.R"))
source(here("R/gene_sets_of_interest.R"))
source(here("R/getSetHeatmap.R"))
source(here("R/venn_diagram_gene_lists.R"))
source(here("R/plotGeneCount.R"))
source(here("R/plot_counts_eds1_rgt1_double.R"))
source(here("R/multiplot.R"))

yeast_gff = rtracklayer::import(Sys.getenv("yeast_gff"))

cc_df = read_csv("../../shared_data/Combined_Eds1_Rgt1_Lys14_CCTargets.csv")

topgo_res = list(
  res_lists = readRDS(here("data/topGO_out/go_results_objects.rds")),
  venn_sets = readRDS(here("data/topGO_out/go_results_objects_by_venn_set.rds"))
)

venn = getVennResults(shrunken_res_lists,1,.05)

long_df_filter_policy = "padj < .05, shrunken lfc > .05"
```

# prep cc for cytoscape
```{r}
x = cc_df %>% pivot_longer(-c(id, symbol), names_to = 'condition', values_to = 'cc_padj') %>% mutate(lysine = ifelse(str_detect(condition, 'plus'), 'plus', 'minus'), ko = str_extract(condition, "\\w+\\d+")) %>% mutate(ko = str_replace('eds1', 'YBR033W')) %>% mutate(ko = str_replace('rgt1', 'YKL038W')) %>% mutate(ko = str_replace('eds1', 'YBR033W'))
=======

source(here("R/create_long_results_df.R"))
source(here("R/gene_sets_of_interest.R"))
source(here("R/get_set_heatmap.R"))
source(here("R/venn_diagram_gene_lists.R"))
source(here("R/plotGeneCount.R"))

yeast_gff = rtracklayer::import(Sys.getenv("yeast_gff"))
>>>>>>> 8fa902069a054d712991ad24f35a0a2007b11ed5
```


# Display design and replicate tally

```{r tally and show design, echo=FALSE}
as_tibble(colData(dds)) %>%
  group_by(genotype, aminoAcid) %>%
  tally() %>%
  arrange(aminoAcid) %>%
  DT::datatable()
```

# cluster significant genes by their profiles
```{r, echo = FALSE}

betas = coef(dds)
colnames(betas)
topGenes <- head(order(results(dds)$padj),20)
mat <- betas[topGenes, -1]

# set maximum LFC to 3 for vis purposes
thr <- 3
mat[abs(mat) < -thr] <- -thr
mat[mat > thr] <- thr

colnames(mat) = str_remove(colnames(mat), "genotype_")
colnames(mat) = str_remove(colnames(mat), "genotype")
colnames(mat) = str_remove(colnames(mat), "aminoAcid")

gene_ids = tibble(id = rownames(mat)) %>%
  left_join(as_tibble(yeast_gff), by = c('id' = 'locus_tag')) %>%
  distinct(id, .keep_all = TRUE)
rownames(mat) = gene_ids$gene

plt = mat %>% 
  as_tibble(mat, rownames = "gene_name") %>% 
  pivot_longer(-gene_name, names_to = "coef", values_to = "effect") %>% 
  mutate(coef = str_replace(coef, "_HisMetLeuUra_vs_LysHisMetLeuUra", "minus_lys")) %>%
  mutate(coef = str_replace(coef, ".HisMetLeuUra", ".minus_lys")) %>%
  mutate(coef = fct_relevel(as.factor(coef),
                            "minus_lys",
                           "EDS1_vs_WT", 
                           "EDS1.minus_lys",
                           "RGT1_vs_WT",
                           "RGT1.minus_lys",
                           "LYS14_vs_WT",
                           "LYS14.minus_lys",
                           "EDS1_RGT1_vs_WT",
                           "EDS1_RGT1.minus_lys",
                           "EDS1_LYS14_vs_WT",
                           "EDS1_LYS14.minus_lys")) %>%
  ggplot + 
  geom_tile(aes(coef, gene_name, fill = effect), color = "black") +
    scale_fill_gradient2(low = "#075AFF",
                         mid = "#FFFFCC",
                         high = "#FF0000") +
    # coord_fixed() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
          text = element_text(size = 20))

  ggsave(here("plots/coefficient_heatmap.png"), plt, device = 'png')

plt

```

# Plot gene with the most significant effect btwn aminoAcid groups

```{r plt gene with most significant btwn group effect, echo = FALSE}
<<<<<<< HEAD
library(foreach)
shrunken_res_lists$compare_interaction$eds1_rgt1_double_vs_single_effects %>%
     as_tibble(rownames = "id") %>%
     filter(abs(log2FoldChange) > 2, padj < .05) %>% 
     left_join(cc_df) %>%
     filter(eds1_plus_lys_pval < .07 | eds1_minus_lys_pval < .07) %>%
  foreach(i = iterators::iter(., by='row')) %do%{
    plotEds1Rgt1DoubleCounts(i$symbol, i$id)
  }
```

```{r}
# re-do this with new venn list
# intersect(gene_lists_for_venn$down$singles$RGT1, gene_lists_for_venn$up$singles$EDS1)
```

# Genes which change in different directions

* There are no genes which change in different directions in EDS1 and RGT1.  
=======
# see https://www.bioconductor.org/packages/devel/workflows/vignettes/rnaseqGene/inst/doc/rnaseqGene.html

interaction_term_res = results(dds)

gene_index = which.min(interaction_term_res$padj) #which(rownames(interaction_term_res) == "YOR105W") #
>>>>>>> 8fa902069a054d712991ad24f35a0a2007b11ed5

* [YPL187W](https://www.yeastgenome.org/locus/S000006108) is up in the EDS1 KO and down in LYS14.

Mating pheromone alpha-factor, made by alpha cells; interacts with mating type a cells to induce cell cycle arrest and other responses leading to mating; also encoded by MF(ALPHA)2, although MF(ALPHA)1 produces most alpha-factor; binds copper(II) ions 3 4 5 6 8 

* [YBR296C, PHO89](https://www.yeastgenome.org/locus/S000000500) is down in EDS1 and 
up in LYS14

Plasma membrane Na+/Pi cotransporter; active in early growth phase; similar to phosphate transporters of Neurospora crassa; transcription regulated by inorganic phosphate concentrations and Pho4p; mutations in related human transporter genes hPit1 and hPit2 are associated with hyperphosphatemia-induced calcification of vascular tissue and familial idiopathic basal ganglia calcification 1 2 3 4 

* there are no genes which change in different directions between RGT1 and LYS14

# Lysine Pathway

Lysine Biosynthesis Genes--Revist
Eds1 activates LYS4, ACO2, LYS9, LYS12, CTP1, LYS2, LYS1, LYS14 in low glucose -lysine.
LYS4 repressed in galactose +/-lysine, ACO2 repressed in galactose -lysine, LYS2 repressed in high glucose.
LYS4, LYS9, ACO2 are the only genes for which |LFC| > 1.

```{r}
getSetHeatMap(long_res_df, gene_set_df_list$lysine_genes, 
              c("EDS1", "RGT1", "EDS1_RGT1", "LYS14", "EDS1_LYS14"), 
              c('plus_lys', 'minus_lys'), paste0("lysine genes -- ", long_df_filter_policy))
```


# same with LYS

# Eds1 has a repressive efect on glyoxylate cycle genes -- combine with TCA network

```{r}
getSetHeatMap(long_res_df, gene_set_df_list$glyoxylate)
```


## Eds1 has a repressive effect on TCA cycle genes

<<<<<<< HEAD
TCA genes should be 'on' in WT galactose. If the TCA genes go up when EDS1 is 
KO'd, then EDS1 does have a repressive effect on TCA genes. Another factor to 
this, however, is that overtime in galactose, we probably expect TCA genes to go 
up.
=======
# ggsave(here("plots/condition_specific_THI4.png"), ptchwrk, device = "pdf")
>>>>>>> 8fa902069a054d712991ad24f35a0a2007b11ed5

```{r}
getSetHeatMap(long_res_df, gene_set_df_list$tca_genes, 
              c("EDS1", "RGT1", "EDS1_RGT1", "LYS14"), 
              c('plus_lys', 'minus_lys'), paste0("tca cycle genes -- ", long_df_filter_policy))
```

<<<<<<< HEAD
```{r}

getSetHeatMap(long_res_df, gene_set_df_list$membrane_transport_genes, 
              c("EDS1", "RGT1", "EDS1_RGT1", "LYS14"), 
              c('plus_lys', 'minus_lys'), 
              paste0("membrane transport -- ", long_df_filter_policy))
```

```{r}

getSetHeatMap(long_res_df, gene_set_df_list$response_to_glucose_genes, 
              c("EDS1", "RGT1", "LYS14", "EDS1_RGT1", "EDS1_LYS14"), 
              c('plus_lys', 'minus_lys'), 
              paste0("response to glucose -- ", long_df_fltr_policy))
```

# Eds1 directly represses SKS1 in both galactose +/- lysine.
SKS1 is a repressor kinase of SNF3 (Yeast Genome), a low glucose sensor required for induction of certain hexose transporters and which triggers inhibition of Rgt1 (Ozcan et al, 1998).

CHASE NOTE: I want to review that SKS1 is a repressor of SNF3. SGD says it is 
indepdendent of the SNF3 pathway, and that SKS1 is required in low gluc in 
SNF3 null strains. Confusing b/c SKS means suppressor kinase of snf3.

 The snf3 genomic suppressors, rgt1, RGT2 and ssn6, are also capable of suppressing snf3 associated growth defects in a strain lacking sks1. 


```{r}
getSetHeatMap_aug = function(long_res_df, pathway_genes_df, select_genotypes,
                         select_conditions, title, text_size = 20){

  long_res_df %>%
    filter(id %in% pathway_genes_df$id) %>%
    mutate(shrunk_lfc = replace_na(shrunk_lfc, 0)) %>%
    # set a cutoff on the LFC to scale the gradation in the heatmap/colors
    # for those genes that are 3 or more, they should be split out and examined
    # separately
    mutate(shrunk_lfc = ifelse(shrunk_lfc > 2, 2, shrunk_lfc)) %>%
    mutate(shrunk_lfc = ifelse(shrunk_lfc < -2, -2, shrunk_lfc)) %>%
    filter(genotype %in% select_genotypes) %>%
    filter(condition %in% select_conditions) %>%
    left_join(pathway_genes_df) %>%
    ggplot() +
    geom_tile(aes(genotype, symbol, fill = shrunk_lfc  ), color = "black") +
    scale_fill_gradient2(low = "#075AFF",
                         mid = "white",
                         high = "#FF0000") +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
          text = element_text(size = text_size)) +
    ggtitle(title) +
    facet_grid(cols = vars(condition))
}

hxt_trns_set = gene_set_df_list$hexose_transmem_trans_genes %>%
  rbind(tibble(symbol = c("HXT1", "HXT2", "HXT12"), id = c("YHR094C", "YMR011W", "YIL170W"))) %>%
  mutate(symbol = factor(symbol, levels = paste0(rep("HXT", 17), seq(17,1))))

getSetHeatMap_aug(long_res_df, hxt_trns_set, 
              c("EDS1", "RGT1", "LYS14", "EDS1_RGT1", "EDS1_LYS14"), 
              c('plus_lys', 'minus_lys'), 
              paste0("hexose tarnsmembrane transporter genes -- ", long_df_filter_policy))
```

```{r}
=======
```{r}
# re-do this with new venn list
# intersect(gene_lists_for_venn$down$singles$RGT1, gene_lists_for_venn$up$singles$EDS1)
```

# Genes which change in different directions

* There are no genes which change in different directions in EDS1 and RGT1.  

* [YPL187W](https://www.yeastgenome.org/locus/S000006108) is up in the EDS1 KO and down in LYS14.

Mating pheromone alpha-factor, made by alpha cells; interacts with mating type a cells to induce cell cycle arrest and other responses leading to mating; also encoded by MF(ALPHA)2, although MF(ALPHA)1 produces most alpha-factor; binds copper(II) ions 3 4 5 6 8 

* [YBR296C, PHO89](https://www.yeastgenome.org/locus/S000000500) is down in EDS1 and 
up in LYS14

Plasma membrane Na+/Pi cotransporter; active in early growth phase; similar to phosphate transporters of Neurospora crassa; transcription regulated by inorganic phosphate concentrations and Pho4p; mutations in related human transporter genes hPit1 and hPit2 are associated with hyperphosphatemia-induced calcification of vascular tissue and familial idiopathic basal ganglia calcification 1 2 3 4 

* there are no genes which change in different directions between RGT1 and LYS14

# Lysine Pathway

Lysine Biosynthesis Genes--Revist
Eds1 activates LYS4, ACO2, LYS9, LYS12, CTP1, LYS2, LYS1, LYS14 in low glucose -lysine.
LYS4 repressed in galactose +/-lysine, ACO2 repressed in galactose -lysine, LYS2 repressed in high glucose.
LYS4, LYS9, ACO2 are the only genes for which |LFC| > 1.

```{r}
getSetHeatMap(long_res_df, gene_set_df_list$lysine_genes, 
              c("EDS1_", "RGT1_", "EDS1_RGT1_", "LYS14_", "EDS1_LYS14_"), 
              c('plus_lys', 'minus_lys'), "lysine genes -- unfiltered for lfc or padj")
```
>>>>>>> 8fa902069a054d712991ad24f35a0a2007b11ed5

getSetHeatMap(long_res_df, gene_set_df_list$glucose_transmem_trans_genes, 
              c("EDS1", "RGT1", "LYS14", "EDS1_RGT1", "EDS1_LYS14"), 
              c('plus_lys', 'minus_lys'), 
              paste0("glucose transmembrane transport genes -- ", long_df_filter_policy))
```

<<<<<<< HEAD
```{r}

getSetHeatMap(long_res_df, gene_set_df_list$glucose_transmem_transport_activity_genes, 
              c("EDS1", "RGT1", "LYS14", "EDS1_RGT1", "EDS1_LYS14"), 
              c('plus_lys', 'minus_lys'), 
              paste0("glucose transmem trans activity -- ", long_df_filter_policy))
```

```{r}
getSetHeatMap(long_res_df, gene_set_df_list$glucose_import_genes, 
              c("EDS1", "RGT1", "LYS14", "EDS1_RGT1", "EDS1_LYS14"), 
              c('plus_lys', 'minus_lys'), 
              paste0("glucose import -- ", long_df_filter_policy))
```

```{r}
getSetHeatMap(long_res_df, gene_set_df_list$hexokinase_genes, 
              c("EDS1", "RGT1", "LYS14", "EDS1_RGT1", "EDS1_LYS14"), 
              c('plus_lys', 'minus_lys'), 
              paste0("hexokinase -- ", long_df_filter_policy))
=======
# same with LYS

# Eds1 has a repressive efect on glyoxylate cycle genes -- combine with TCA network

## Eds1 has a repressive effect on TCA cycle genes

TCA genes should be 'on' in WT galactose. If the TCA genes go up when EDS1 is 
KO'd, then EDS1 does have a repressive effect on TCA genes. Another factor to 
this, however, is that overtime in galactose, we probably expect TCA genes to go 
up.

```{r}
getSetHeatMap(long_res_df, gene_set_df_list$tca_genes, 
              c("EDS1_", "RGT1_", "EDS1_RGT1_", "LYS14_"), 
              c('plus_lys', 'minus_lys'), "tca cycle genes -- unfiltered for lfc or padj")
>>>>>>> 8fa902069a054d712991ad24f35a0a2007b11ed5
```

```{r}

getSetHeatMap(long_res_df, gene_set_df_list$membrane_transport_genes, 
              c("EDS1_", "RGT1_", "EDS1_RGT1_", "LYS14_"), 
              c('plus_lys', 'minus_lys'), 
              "membrane transport -- unfiltered for lfc or padj")
```


```{r}

<<<<<<< HEAD
getSetHeatMap(long_res_df, gene_set_df_list$glyoxylate_genes, 
              c("EDS1", "RGT1", "EDS1_RGT1", "LYS14"), 
              c('plus_lys', 'minus_lys'), paste0("glyoxylate genes -- ", long_df_filter_policy))

```

# Fermentation pathways?

These should be 'off' in galactose. If perturbing the genes in question cause 
fermentation pathway genes to increase, then it is evidence that these genes 
have a role in fermentation

```{r}
getSetHeatMap(long_res_df, gene_set_df_list$anaerobic_resp_genes, 
              c("EDS1", "RGT1", "EDS1_RGT1", "LYS14"), 
              c('plus_lys', 'minus_lys'), paste0("anaerobic respiration -- ", long_df_filter_policy))
```


# Aerobic respiration pathways

These should be 'on' in galactose, presumably. Down regulation would point to 
genes that a given perturbed locus has a positive impact on. up regulation points 
to genes which a given perturbed locus is supressing. Expecting few of those?

```{r}
getSetHeatMap(long_res_df, gene_set_df_list$aerobic_resp_genes, 
              c("EDS1", "RGT1", "EDS1_RGT1", "LYS14"), 
              c('plus_lys', 'minus_lys'), paste0("aerobic respiration -- ", long_df_filter_policy))
```

=======
getSetHeatMap(long_res_df, gene_set_df_list$response_to_glucose_genes, 
              c("EDS1_", "RGT1_", "EDS1_RGT1_", "LYS14_"), 
              c('plus_lys', 'minus_lys'), 
              "response to glucose -- unfiltered for lfc or padj")
```

# Eds1 directly represses SKS1 in both galactose +/- lysine.
SKS1 is a repressor kinase of SNF3 (Yeast Genome), a low glucose sensor required for induction of certain hexose transporters and which triggers inhibition of Rgt1 (Ozcan et al, 1998).

CHASE NOTE: I want to review that SKS1 is a repressor of SNF3. SGD says it is 
indepdendent of the SNF3 pathway, and that SKS1 is required in low gluc in 
SNF3 null strains. Confusing b/c SKS means suppressor kinase of snf3.

 The snf3 genomic suppressors, rgt1, RGT2 and ssn6, are also capable of suppressing snf3 associated growth defects in a strain lacking sks1. 


```{r}

getSetHeatMap(long_res_df, gene_set_df_list$hexose_transmem_trans_genes, 
              c("EDS1_", "RGT1_", "EDS1_RGT1_", "LYS14_"), 
              c('plus_lys', 'minus_lys'), 
              "hexose tarnsmembrane transporter genes -- unfiltered for lfc or padj")
```

```{r}

getSetHeatMap(long_res_df, gene_set_df_list$glucose_transmem_trans_genes, 
              c("EDS1_", "RGT1_", "EDS1_RGT1_", "LYS14_"), 
              c('plus_lys', 'minus_lys'), 
              "glucose transmembrane transport genes -- unfiltered for lfc or padj")
```

```{r}

getSetHeatMap(long_res_df, gene_set_df_list$glucose_transmem_transport_activity_genes, 
              c("EDS1_", "RGT1_", "EDS1_RGT1_", "LYS14_"), 
              c('plus_lys', 'minus_lys'), 
              "glucose transmem trans activity -- unfiltered for lfc or padj")
```
>>>>>>> 8fa902069a054d712991ad24f35a0a2007b11ed5

```{r}
getSetHeatMap(long_res_df, gene_set_df_list$glucose_import_genes, 
              c("EDS1_", "RGT1_", "EDS1_RGT1_", "LYS14_"), 
              c('plus_lys', 'minus_lys'), 
              "glucose import -- unfiltered for lfc or padj")
```

<<<<<<< HEAD
getSetHeatMap(long_res_df, gene_set_df_list$glycolytic_genes, 
              c("EDS1", "RGT1", "EDS1_RGT1", "LYS14"), 
              c('plus_lys', 'minus_lys'), paste0("glycolysis genes -- ", long_df_filter_policy))

```

* Eds1 strongly and directly activates HXK1, where Rgt1 does not.

* Eds1 directly activates CDC19 in galactose +lysine.

CDC19 synthesizes pyruvate from phosphoenolpyruvate, the last step before TCA.
Rgt1 directly represses CDC19 in our -lysine conditions, but activates in galactose +lysine.

CHASE QUESTION: why would the presence/absence of lysine affect TCA cycle?
=======
```{r}
getSetHeatMap(long_res_df, gene_set_df_list$hexokinase_genes, 
              c("EDS1_", "RGT1_", "EDS1_RGT1_", "LYS14_"), 
              c('plus_lys', 'minus_lys'), 
              "hexokinase -- unfiltered for lfc or padj")
```



```{r}

getSetHeatMap(long_res_df, gene_set_df_list$glyoxylate_genes, 
              c("EDS1_", "RGT1_", "EDS1_RGT1_", "LYS14_"), 
              c('plus_lys', 'minus_lys'), "glyoxylate genes -- unfiltered for lfc or padj")

```
>>>>>>> 8fa902069a054d712991ad24f35a0a2007b11ed5

# Eds1 and Rgt1 directly regulate many common targets in the same direction

<<<<<<< HEAD
```{r}
png(here("plots/plus_lysine_venn.png"),width = 800, height = 800, bg = "white")
grid.arrange(venn$plots$plus_lys$up$singles, 
          venn$plots$plus_lys$down$singles,
          venn$plots$minus_lys$up$singles, 
          venn$plots$minus_lys$down$singles,
          top = grid::textGrob("Plus Lysine",gp=gpar(fontsize=30,font=3)),
          bottom = grid::textGrob("Minus Lysine",gp=gpar(fontsize=30,font=3)))
dev.off()

grid.arrange(venn$plots$plus_lys$up$singles, 
          venn$plots$plus_lys$down$singles,
          venn$plots$minus_lys$up$singles, 
          venn$plots$minus_lys$down$singles,
          top = textGrob("Plus Lysine",gp=gpar(fontsize=30,font=3)),
          bottom = textGrob("Minus Lysine",gp=gpar(fontsize=30,font=3)))
```

```{r}
DT::datatable(topgo_res$venn_sets$bp$plus_lys$eds1_only$results$summary, 
              class = "display")
```

```{r}
library(gridExtra)
grid.arrange(venn$plots$plus_lys$up$eds1_vs_lys14, 
          venn$plots$plus_lys$down$eds1_vs_lys14,
          venn$plots$minus_lys$up$eds1_vs_lys14, 
          venn$plots$minus_lys$down$eds1_vs_lys14,
          top = textGrob("Plus Lysine",gp=gpar(fontsize=30,font=3)),
          bottom = textGrob("Minus Lysine",gp=gpar(fontsize=30,font=3)))
```

```{r}
grid.arrange(venn$plots$plus_lys$up$eds1_vs_rgt1, 
          venn$plots$plus_lys$down$eds1_vs_rgt1,
          venn$plots$minus_lys$up$eds1_vs_rgt1, 
          venn$plots$minus_lys$down$eds1_vs_rgt1,
          top = textGrob("Plus Lysine",gp=gpar(fontsize=30,font=3)),
          bottom = textGrob("Minus Lysine",gp=gpar(fontsize=30,font=3)))
```

Do the same with complements

# Reguation of Eds1 by RGT1 or LYS14?

```{r}

bind_rows(shrunken_res_lists$plus_lys$LYS14 %>%
              as_tibble(rownames = 'gene') %>%
              select(gene, log2FoldChange, padj) %>%
              mutate(condition = 'plus_lys',
                     genotype = 'LYS14'),
            shrunken_res_lists$minus_lys$LYS14 %>%
              as_tibble(rownames = 'gene') %>%
              select(gene, log2FoldChange, padj) %>%
              mutate(condition = 'minus_lys',
                     genotype = 'LYS14')
            ) %>%
  filter(gene %in% c('YKL038W', 'YBR033W')) %>%
  ggplot() + 
  geom_bar(aes(interaction(condition, genotype), log2FoldChange, fill = gene), stat='identity', position = 'dodge') + ylim(-.5, .5)

```

```{r}

bind_rows(shrunken_res_lists$plus_lys$RGT1 %>%
=======
These should be 'off' in galactose. If perturbing the genes in question cause 
fermentation pathway genes to increase, then it is evidence that these genes 
have a role in fermentation

```{r}
getSetHeatMap(long_res_df, gene_set_df_list$anaerobic_resp_genes, 
              c("EDS1_", "RGT1_", "EDS1_RGT1_", "LYS14_"), 
              c('plus_lys', 'minus_lys'), "anaerobic respiration -- unfiltered for lfc or padj")
```


# Aerobic respiration pathways

These should be 'on' in galactose, presumably. Down regulation would point to 
genes that a given perturbed locus has a positive impact on. up regulation points 
to genes which a given perturbed locus is supressing. Expecting few of those?

```{r}
getSetHeatMap(long_res_df, gene_set_df_list$aerobic_resp_genes, 
              c("EDS1_", "RGT1_", "EDS1_RGT1_", "LYS14_"), 
              c('plus_lys', 'minus_lys'), "aerobic respiration -- unfiltered for lfc or padj")
```


```{r}

getSetHeatMap(long_res_df, gene_set_df_list$glycolytic_genes, 
              c("EDS1_", "RGT1_", "EDS1_RGT1_", "LYS14_"), 
              c('plus_lys', 'minus_lys'), "glycolysis genes -- unfiltered for lfc or padj")

```

* Eds1 strongly and directly activates HXK1, where Rgt1 does not.

* Eds1 directly activates CDC19 in galactose +lysine.

CDC19 synthesizes pyruvate from phosphoenolpyruvate, the last step before TCA.
Rgt1 directly represses CDC19 in our -lysine conditions, but activates in galactose +lysine.

CHASE QUESTION: why would the presence/absence of lysine affect TCA cycle?

# Eds1 and Rgt1 directly regulate many common targets in the same direction

```{r}
venn$plots$minus_lys$up$singles
```



```{r}
venn$plots$minus_lys$down$singles
```

```{r}
venn$plots$plus_lys$up$singles
```

```{r}
venn$plots$plus_lys$down$singles
```


intersect top up/down genes in eds1/rgt1, do GO

Do the same with complements

# Reguation of Eds1 by RGT1 or LYS14?

```{r}

bind_rows(shrunken_res_lists$plus_lys$LYS14 %>%
>>>>>>> 8fa902069a054d712991ad24f35a0a2007b11ed5
              as_tibble(rownames = 'gene') %>%
              select(gene, log2FoldChange, padj) %>%
              mutate(condition = 'plus_lys',
                     genotype = 'LYS14'),
<<<<<<< HEAD
            shrunken_res_lists$minus_lys$RGT1 %>%
=======
            shrunken_res_lists$minus_lys$LYS14 %>%
>>>>>>> 8fa902069a054d712991ad24f35a0a2007b11ed5
              as_tibble(rownames = 'gene') %>%
              select(gene, log2FoldChange, padj) %>%
              mutate(condition = 'minus_lys',
                     genotype = 'LYS14')
            ) %>%
<<<<<<< HEAD
  filter(gene %in% c('YDR034C', 'YBR033W')) %>%
  ggplot() + 
  geom_bar(aes(interaction(condition, genotype), log2FoldChange, fill = gene), stat='identity', position = 'dodge') + ylim(-.5, .5)


```


=======
  filter(gene %in% c('YKL038W', 'YBR033W')) %>%
  ggplot() + 
  geom_bar(aes(interaction(condition, genotype), log2FoldChange, fill = gene), stat='identity', position = 'dodge') + ylim(-.5, .5)

```

>>>>>>> 8fa902069a054d712991ad24f35a0a2007b11ed5
```{r}

dds_vst = varianceStabilizingTransformation(dds, blind=FALSE)

assays(dds_vst)[[1]] %>%
  as_tibble(rownames = 'gene') %>% 
  pivot_longer(-gene, names_to = 'sample', values_to = 'vst') %>%
  left_join(as_tibble(colData(dds_vst))) %>%
  filter(gene %in% c('YKL038W', 'YBR033W'), genotype %in% c('WT', "LYS14")) %>%
  ggplot() + 
  geom_boxplot(aes(interaction(genotype, aminoAcid), vst, fill = gene)) + 
  geom_point(aes(interaction(genotype, aminoAcid), vst))
  

```


# HXT expression

heatmap of all HXT genes

# RGT1/MIG pathway genes

heatmap of these genes

Eds1 directly represses MTH1 in galactose with lysine.
  This needs binding data

  Rgt1 represses in galactose with lysine, galactose without lysine, high glucose late.
  
  Mig1 represses in high glucose early, high glucose late.
  Mig2 activates in high glucose early, represses in high glucose late.
  Rgt1, Mig1, Mig2 all represses MTH1 according to Kuttykrishnan et al.
  Eds1 indirectly represses STD1 in galactose with lysine and galactose without lysine.
  Rgt1 represses STD1 in galactose with lysine, galactose without lysine, and low glucose.
  Eds1 directly represses MIG1 in galactose no lysine, activates in glucose.
  Rgt1 represses MIG1 in galactose +/- lysine.

```{r}
plotGeneCountCustom(dds, 'YBR033W', 'EDS1')$combined
```

```{r}
plotGeneCountCustom(dds, 'YKL038W', 'RGT1')$combined
```

```{r}
plotGeneCountCustom(dds, 'YGL035C', 'MIG1')$combined
```

```{r}
plotGeneCountCustom(dds, 'YGL209W', 'MIG2')$combined
```
<<<<<<< HEAD
=======

```{r}
plotGeneCountCustom(dds, 'YGL209W', 'MIG2')$combined
```
>>>>>>> 8fa902069a054d712991ad24f35a0a2007b11ed5

# Phosphate Regulation

* Eds1 causes highly significant repression of PHO11, PHO12, PHO89 (11-fold), SPL2 (32-fold) across all time points.
* Note: not so true in galac, though there is a down reg and what seems to be an 
  additive effect
* no evidence of direct binding

* Rgt1 also causes highly significant repression of these genes in glucose -lysine. 
  remains to be checked

```{r}
plotGeneCountCustom(dds, 'YBR296C', 'PHO89')$combined
```
```{r}
plotGeneCountCustom(dds, 'YAR071W', 'PHO11')$combined
```
```{r}
plotGeneCountCustom(dds, 'YHR215W', 'PHO12')$combined
```
```{r}
plotGeneCountCustom(dds, 'YHR136C', 'SPL2')$combined
```

```{r}
getSetHeatMap(long_res_df, gene_set_df_list$acid_phosphate_activity, 
<<<<<<< HEAD
              c("EDS1", "RGT1", "EDS1_RGT1", "LYS14"), 
=======
              c("EDS1_", "RGT1_", "EDS1_RGT1_", "LYS14_"), 
>>>>>>> 8fa902069a054d712991ad24f35a0a2007b11ed5
              c('plus_lys', 'minus_lys'), "acid phosphate activity -- unfiltered for lfc or padj")
```

```{r}
getSetHeatMap(long_res_df, gene_set_df_list$phosphate_ion_transmem_transport, 
<<<<<<< HEAD
              c("EDS1", "RGT1", "EDS1_RGT1", "LYS14"), 
=======
              c("EDS1_", "RGT1_", "EDS1_RGT1_", "LYS14_"), 
>>>>>>> 8fa902069a054d712991ad24f35a0a2007b11ed5
              c('plus_lys', 'minus_lys'), "phosphate_ion_transmem_transport -- unfiltered for lfc or padj")
```

```{r}
getSetHeatMap(long_res_df, gene_set_df_list$phoscontaining_compound_meta_proc, 
<<<<<<< HEAD
              c("EDS1", "RGT1", "EDS1_RGT1", "LYS14"), 
=======
              c("EDS1_", "RGT1_", "EDS1_RGT1_", "LYS14_"), 
>>>>>>> 8fa902069a054d712991ad24f35a0a2007b11ed5
              c('plus_lys', 'minus_lys'), "phoscontaining_compound_meta_proc -- unfiltered for lfc or padj")
```


# Transcription factors
<<<<<<< HEAD
Eds1 directly activates Mig3 in high glucose -lysine.

```{r}
plotGeneCountCustom(dds, 'YER028C', 'MIG3')$combined
```

Eds1 directly represses MIG1 in galactose no lysine, activates in glucose.
(see plot above)

Rgt1 represses MIG1 in galactose +/- lysine.
(see plot above)

Eds1 directly represses MTH1 in galactose with lysine.
```{r}
plotGeneCountCustom(dds, 'YDR277C', 'MTH1')$combined
```

```{r}
plotGeneCountCustom(dds, 'YIL074C', 'SER33')$combined
```

# Binding Data

```{r}
library(ggmsa)


sig_binding_eds1 = cc_df %>% 
  filter(eds1_minus_lys_pval < .00005, eds1_plus_lys_pval < .00005) %>%
  pull(id)

library(TxDb.Scerevisiae.UCSC.sacCer3.sgdGene)
library(BSgenome.Scerevisiae.UCSC.sacCer3)

yeast_gff = TxDb.Scerevisiae.UCSC.sacCer3.sgdGene
yeast_genes = genes(yeast_gff)

yeast_genome = BSgenome.Scerevisiae.UCSC.sacCer3

eds1_bound_genes = yeast_genes[yeast_genes$gene_id %in% sig_binding_eds1]

eds1_bound_genes_promoters = promoters(eds1_bound_genes, upstream =150, downstream = 0)

eds1_bound_gene_prom_seqs = getSeq(yeast_genome, eds1_bound_genes_promoters)

pfm <- consensusMatrix(eds1_bound_gene_prom_seqs, baseOnly = TRUE)
pwm <- PWM(consensusMatrix(eds1_bound_gene_prom_seqs))

ggmsa(eds1_bound_gene_prom_seqs, 50, 100, seq_name = TRUE, color = "Chemistry_NT") + geom_seqlogo() + geom_msaBar()

```

```{r}
colnames(cc_df) = tolower(colnames(cc_df))

goi_list = list(
  hexose_trns_and_regs = c(
    'HXT1', 
    'HXT2', 
    'HXT3', 
    'HXT4', 
    'HXT6', 
    'HXT7',
    'MIG2', 
    'MIG3', 
    'MTH1',
    'STD1'),
  glucose_to_citrate = c(
    'CDC19',
    'HXK1',
    'HXK2',
    'GLK1',
    'UCC1'
  ),
  ribosome_biogen = c(
    'RPL3',
    'HAS1'
  ),
  oxidative_or_dna_rep_stress = c(
    'AQR1',
    'AHP1',
    'DAK2',
    'ECM4',
    'RMI1',
    'TOP3',
    'SGS1',
    'GTO1',
    'GLN1',
    'TRR1'
  ),
  aa_transport = c(
    'Uga4',
    'Atg22',
    'Vba1',
    'Vba2',
    'Vba3',
    'Avt1',
    'Avt3',
    'Avt4',
    'Avt6',
    'Avt7',
    'Ers1',
    'Ypq1',
    'Ypq2',
    'Ypq3',
    'Aqr1',
    'QDR1',
    'Mmp1',
    'Vba5',
    'Yct1',
    'Gap1',
    'Agp1',
    'Bap2',
    'Bap3',
    'Agp2',
    'Agp3',
    'Gnp1',
    'Tat1',
    'Tat2',
    'Tat3',
    'Dip5',
    'Put4',
    'Alp1',
    'Can1',
    'Hip1',
    'Lyp1',
    'Mup1',
    'Mup3'
  )
)

goi_list$aa_transport = toupper(goi_list$aa_transport)

x = reduce(list(
  as_tibble(shrunken_res_lists$plus_lys$EDS1, rownames = 'id') %>%
    rename_at(vars(!starts_with("id")), ~paste0( ., "_eds1_plus_lys")),
  as_tibble(shrunken_res_lists$minus_lys$EDS1, rownames = 'id')%>%
    rename_at(vars(!starts_with("id")), ~paste0( ., "_eds1_minus_lys")),
  as_tibble(shrunken_res_lists$plus_lys$RGT1, rownames = 'id')%>%
    rename_at(vars(!starts_with("id")), ~paste0( ., "_rgt1_plus_lys")),
  as_tibble(shrunken_res_lists$minus_lys$RGT1, rownames = 'id')%>%
    rename_at(vars(!starts_with("id")), ~paste0( ., "_rgt1_minus_lys")), 
  as_tibble(shrunken_res_lists$plus_lys$LYS14, rownames = 'id')%>%
    rename_at(vars(!starts_with("id")), ~paste0( ., "_lys14_plus_lys")),
  as_tibble(shrunken_res_lists$minus_lys$LYS14, rownames = 'id')%>%
    rename_at(vars(!starts_with("id")), ~paste0( ., "_lys14_minus_lys")), 
  as_tibble(shrunken_res_lists$plus_lys$EDS1_RGT1, rownames = 'id')%>%
    rename_at(vars(!starts_with("id")), ~paste0( ., "_eds1_rgt1_plus_lys")),
  as_tibble(shrunken_res_lists$minus_lys$EDS1_RGT1, rownames = 'id')%>%
    rename_at(vars(!starts_with("id")), ~paste0( ., "_eds1_rgt1_minus_lys")), 
  as_tibble(shrunken_res_lists$plus_lys$EDS1_LYS14, rownames = 'id')%>%
    rename_at(vars(!starts_with("id")), ~paste0( ., "_eds1_lys14_plus_lys")),
  as_tibble(shrunken_res_lists$minus_lys$EDS1_LYS14, rownames = 'id')%>%
    rename_at(vars(!starts_with("id")), ~paste0( ., "_eds1_lys14_minus_lys"))
), left_join, by = "id") %>%
  pivot_longer(-id) %>%
  mutate(metric = str_extract(name, '(.*?)\\_')) %>%
  mutate(name = str_remove(name, '(.*?)\\_'),
         metric = str_remove(metric, "\\_")) %>%
  mutate(ko = str_extract(name, '(.*?)\\_plus|(.*?)\\_minus')) %>%
  mutate(ko = str_remove(ko, "_plus|_minus")) %>%
  mutate(lysine = ifelse(str_detect(name, "plus"), 'plus', 'minus')) %>%
  dplyr::select(-name) %>%
  pivot_wider(names_from = metric) %>%
  left_join(cc_df %>% 
              pivot_longer(-c(id,symbol), values_to = 'cc_pval') %>% 
              mutate(lysine = 
                       ifelse(str_detect(name, 'plus'), 
                              'plus', 
                              'minus'),
                     ko = str_extract(name, '(.*?)\\_')) %>%
              mutate(ko = str_remove(ko, "_$")) %>%
              dplyr::select(-name)) %>%
  filter(complete.cases(.))
```


```{r}
plt = x %>%
  filter(symbol %in% goi_list$glucose_to_citrate) %>%
  mutate(cc_pval = -log10(cc_pval)) %>%
  mutate(cc_pval = ifelse(cc_pval == Inf, 10, cc_pval)) %>%
  ggplot(aes(log2FoldChange,cc_pval)) +
  geom_point(aes(color = symbol)) +
  geom_hline(yintercept = -log10(.05)) +
  geom_vline(xintercept = .5) +
  geom_vline(xintercept = -.5) +
  facet_grid(cols = vars(ko), rows = vars(lysine))

  plotly::ggplotly(plt)
```
aa transport genes
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7405077/#!po=5.17241
```{r}

plt = x %>%
  filter(symbol %in% goi_list$aa_transport) %>%
  mutate(cc_pval = -log10(cc_pval)) %>%
  mutate(cc_pval = ifelse(cc_pval == Inf, 10, cc_pval)) %>%
  ggplot(aes(log2FoldChange,cc_pval)) +
  geom_point(aes(color = symbol)) +
  geom_hline(yintercept = -log10(.05)) +
  geom_vline(xintercept = .5) +
  geom_vline(xintercept = -.5) +
  facet_grid(cols = vars(ko), rows = vars(lysine)) +
  ggtitle("AA transport")


  plotly::ggplotly(plt)
```

```{r}
plt = x %>%
  filter(symbol %in% c("SOD1", "SOD2"), padj < .05) %>%
  mutate(cc_pval = -log10(cc_pval)) %>%
  mutate(cc_pval = ifelse(cc_pval == Inf, 10, cc_pval)) %>%
  ggplot(aes(log2FoldChange,cc_pval)) +
  geom_point(aes(color = symbol)) +
  geom_hline(yintercept = -log10(.05)) +
  geom_vline(xintercept = .5) +
  geom_vline(xintercept = -.5) +
  facet_grid(cols = vars(ko), rows = vars(lysine))

  plotly::ggplotly(plt)
```

```{r}

plt = x %>%
  filter(symbol %in% c("HXT9", "HXT11")) %>%
  mutate(cc_pval = -log10(cc_pval)) %>%
  mutate(cc_pval = ifelse(cc_pval == Inf, 10, cc_pval)) %>%
  ggplot(aes(log2FoldChange,cc_pval)) +
  geom_point(aes(color = symbol)) +
  geom_hline(yintercept = -log10(.05)) +
  geom_vline(xintercept = .5) +
  geom_vline(xintercept = -.5) +
  facet_grid(cols = vars(ko), rows = vars(lysine))

  plotly::ggplotly(plt)
```


```{r}
plt = x %>%
  filter(symbol %in% goi_list$ribosome_biogen) %>%
  mutate(cc_pval = -log10(cc_pval)) %>%
  mutate(cc_pval = ifelse(cc_pval == Inf, 10, cc_pval)) %>%
  ggplot(aes(log2FoldChange,cc_pval)) +
  geom_point(aes(color = symbol)) +
  geom_hline(yintercept = -log10(.05)) +
  geom_vline(xintercept = .5) +
  geom_vline(xintercept = -.5) +
  facet_grid(cols = vars(ko), rows = vars(lysine))

  plotly::ggplotly(plt)
```

```{r}

plt = x %>%
  filter(symbol %in% goi_list$oxidative_or_dna_rep_stress) %>%
  mutate(cc_pval = -log10(cc_pval)) %>%
  mutate(cc_pval = ifelse(cc_pval == Inf, 10, cc_pval)) %>%
  ggplot(aes(log2FoldChange,cc_pval)) +
  geom_point(aes(color = symbol)) +
  geom_hline(yintercept = -log10(.05)) +
  geom_vline(xintercept = .5) +
  geom_vline(xintercept = -.5) +
  facet_grid(cols = vars(ko), rows = vars(lysine))

  plotly::ggplotly(plt)

```


```{r}

plt = x %>%
  filter(cc_pval < .05, log2FoldChange > 1) %>%
  mutate(cc_pval = -log10(cc_pval)) %>%
  mutate(cc_pval = ifelse(cc_pval == Inf, 10, cc_pval)) %>%
  ggplot(aes(log2FoldChange,cc_pval)) +
  geom_point(aes(color = symbol)) +
  geom_hline(yintercept = -log10(.05)) +
  geom_vline(xintercept = .5) +
  geom_vline(xintercept = -.5) +
  facet_grid(cols = vars(ko), rows = vars(lysine))

  plotly::ggplotly(plt)

```

# Effect of EDS1 on lysine gene expr

```{r}

dds_vst = vst(dds, blind = FALSE)

library(foreach)
library(doParallel)
lysine_expr = foreach(
  i = gene_set_df_list$lysine_genes$id,
  .combine = 'rbind'
) %do% {
  x = plotCounts(dds, 
             gene = i, 
             intgroup = c("aminoAcid", "genotype"), 
             returnData = TRUE,
             normalized = TRUE,
             transform = TRUE)
  x$gene = gene_set_df_list$lysine_genes %>%
    filter(id == i) %>%
    pull(symbol)
  x
}

lysine_expr %>%
  as_tibble() %>%
  filter(genotype %in% c('EDS1', 'WT')) %>%
  group_by(gene,genotype, aminoAcid) %>%
  summarize(med_count = median(count)) %>%
  pivot_wider(names_from = c(genotype, aminoAcid), values_from = med_count) %>%
  dplyr::select(WT_plus_lys, WT_minus_lys, EDS1_minus_lys) %>%
  pivot_longer(-c(gene), names_to = 'cond', values_to = 'count') %>%
  mutate(lys = ifelse(str_detect(cond, 'minus'), 'minus_lys', 'plus_lys')) %>%
  mutate(cond = str_remove(cond, "_plus_lys|_minus_lys")) %>%
  mutate(lys = factor(lys, levels = c("plus_lys", "minus_lys"))) -> data

  ggplot(data, aes(x=lys, y=count, color = gene, group = gene)) +
  scale_y_log10() + 
  geom_point(size = 4) +
  geom_line(data = filter(data, (cond == 'EDS1' | (cond == 'WT' & lys == 'plus_lys'))),
            aes(x = lys, y = count, group = gene), color = "red") +
  geom_line(data = filter(data, (cond == 'WT' | (cond == 'WT' & lys == 'plus_lys'))),
            aes(x = lys, y = count, group = gene),  color = "black")
    
```

```{r}
lysine_expr %>%
  as_tibble() %>%
  filter(genotype %in% c('EDS1', 'WT')) %>%
  group_by(gene,genotype, aminoAcid) %>%
  summarize(med_count = median(count)) %>%
  filter(gene %in% c('ARO9', 'ACO2')) %>%
  ggplot(aes(x=aminoAcid, y=med_count, color=gene, group=gene)) +
  scale_y_log10() + 
  geom_point() +
  geom_line() +
  facet_grid(vars(genotype))
```

```{r}
library(ggsci)
x %>%
  filter(symbol %in% c('HXT2', 'HXT4', 'HXT6', 'HXT7', 'MTH1'), cc_pval < .05) %>%
  mutate(log2FoldChange = ifelse(padj > .05, 0, log2FoldChange)) %>%
  mutate(symbol = factor(symbol, 
                         levels = goi_list$hexose_trns_and_regs)) %>%
  droplevels() %>%
  ggplot() + 
  geom_bar(aes(symbol, log2FoldChange, 
               fill=ko), 
           stat='identity', 
           position = 'dodge',
           width = .4) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_fill_manual(values = c("rgt1" = pal_npg("nrc")(10)[6], 
                               "eds1" = "orange")) +
  facet_wrap(~lysine, nrow = 2, ncol = 1, strip.position = 'right') +
  ggtitle("High Affinity Hexose Transporters")
```
```{r}
plt = x %>%
  filter(symbol %in% c('HXT2', 'HXT4', 'HXT6', 'HXT7', 'MTH1')) %>%
  mutate(cc_pval = -log10(cc_pval)) %>%
  mutate(cc_pval = ifelse(cc_pval == Inf, 10, cc_pval)) %>%
  ggplot(aes(log2FoldChange,cc_pval)) +
  geom_point(aes(color = symbol)) +
  geom_hline(yintercept = -log10(.05)) +
  geom_vline(xintercept = .5) +
  geom_vline(xintercept = -.5) +
  facet_grid(cols = vars(ko), rows = vars(lysine)) +
  ggtitle("High Affinity Glucose Transporters")

  plotly::ggplotly(plt)
```

```{r}
library(ggsci)
x %>%
  filter(symbol %in% c('HXT1', 'HXT3', 'MIG2', 'STD1'), cc_pval < .05) %>%
  mutate(log2FoldChange = ifelse(padj > .05, 0, log2FoldChange)) %>%
  mutate(symbol = factor(symbol, 
                         levels = goi_list$hexose_trns_and_regs)) %>%
  droplevels() %>%
  ggplot() + 
  geom_bar(aes(symbol, log2FoldChange, 
               fill=ko), 
           stat='identity', 
           position = 'dodge',
           width = .4) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_fill_manual(values = c("rgt1" = pal_npg("nrc")(10)[6], 
                               "eds1" = "orange")) +
  facet_wrap(~lysine, nrow = 2, ncol = 1, strip.position = 'right') +
  ggtitle("Low Affinity Hexose Transporters")
```
```{r}
plt = x %>%
  filter(symbol %in% c('HXT1', 'HXT3', 'MIG2', 'STD1')) %>%
  mutate(cc_pval = -log10(cc_pval)) %>%
  mutate(cc_pval = ifelse(cc_pval == Inf, 10, cc_pval)) %>%
  ggplot(aes(log2FoldChange,cc_pval)) +
  geom_point(aes(color = symbol)) +
  geom_hline(yintercept = -log10(.05)) +
  geom_vline(xintercept = .5) +
  geom_vline(xintercept = -.5) +
  facet_grid(cols = vars(ko), rows = vars(lysine)) +
  ggtitle("Low Affinity Glucose Transporters")

  plotly::ggplotly(plt)
```

```{r}
plt = x %>%
  filter(symbol %in% c('IDH1', 'IDH2', 'ABF1'), padj < .05) %>%
  mutate(cc_pval = -log10(cc_pval)) %>%
  mutate(cc_pval = ifelse(cc_pval == Inf, 10, cc_pval)) %>%
  ggplot(aes(log2FoldChange,cc_pval)) +
  geom_point(aes(color = symbol)) +
  geom_hline(yintercept = -log10(.05)) +
  geom_vline(xintercept = .5) +
  geom_vline(xintercept = -.5) +
  facet_grid(cols = vars(ko), rows = vars(lysine)) +
  ggtitle("Low Affinity Glucose Transporters") +
  theme(axis.text = element_text(size = 20))

  plotly::ggplotly(plt)
```

```{r}
x = long_res_df %>% 
  filter(genotype %in% c("EDS1", "RGT1"), abs(shrunk_lfc) > .5) %>% 
  left_join(cc_df) %>% 
  filter(eds1_plus_lys_pval < .1 | 
           eds1_minus_lys_pval < .1 | 
           rgt1_plus_lys_pval < .1 | 
           rgt1_minus_lys_pval < .1) %>% 
  group_by(genotype) %>% 
  group_split()
names(x) = c('EDS1', 'RGT1')

eds1_only = setdiff(pull(filter(x$EDS1, eds1_plus_lys_pval < .1 | 
           eds1_minus_lys_pval < .1), symbol), pull(filter(x$RGT1, rgt1_plus_lys_pval < .1 | 
           rgt1_minus_lys_pval < .1), symbol)) 

x$RGT1 %>% 
  filter(id %in% go2gn$`GO:0019319`) %>% View()


```

```{r}
symbol_vec = select(org.Sc.sgd.db, keys = go2gn$`GO:0019319`, keytype = 'ORF', columns = c('COMMON', 'DESCRIPTION'))

symbol_vec_df = symbol_vec %>%
  as_tibble() %>%
  distinct(ORF, .keep_all = TRUE)

hexose_biosyn_df = tibble(
  tibble(id = go2gn$`GO:0019319`)
) %>% left_join(symbol_vec_df, by = c('id' = 'ORF')) %>%
  dplyr::rename(symbol = COMMON) %>%
  distinct(id, .keep_all = TRUE)

hexose_biosynthesis = getSetHeatMap(
  long_res_df,hexose_biosyn_df, c("EDS1", "RGT1", "LYS14", "EDS1_RGT1"), c("plus_lys", "minus_lys"), 'hexose biosynthetic'
)
```


```{r}
x %>%
  filter(id %in% gene_set_df_list$lysine_genes$id) %>%
  mutate(symbol = factor(symbol, levels = c("ACO2", "ARO8", "ARO9",
                                            'LYS1', 'LYS2', 'LYS4', 'LYS5',
                                            'LYS9', 'LYS12', 'LYS20', 'LYS21'))) %>%
  ggplot() + 
  geom_bar(aes(symbol, log2FoldChange, 
               fill=ko), stat='identity', position = 'dodge') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_fill_manual(values = c("rgt1" = pal_npg("nrc")(10)[6], 
                               "eds1" = "orange", 
                               "lys14" = pal_npg("nrc")(10)[1])) +
  facet_wrap(~lysine, nrow = 2, ncol = 1, strip.position = 'right')
```

```{r}
map2(gene_set_df_list$lysine_genes$id, gene_set_df_list$lysine_genes$symbol,
     ~plotGeneCountCustom(dds_lrt, .x, .y)$combined)
```

```{r}
glucose_upreg_genes = c(
  'YLR420W',
'YEL040W',
'YOR047C',
'YER028C',
'YKL068W-A',
'YCR021C',
'YPR156C',
'YOR107W',
'YBL069W',
'YNCP0005C',
'YBR085W',
'YBR238C',
'YKL216W',
'YOL014W',
'YNR060W',
'YNL024C',
'YLR413W',
'YDL241W',
'YDL039C',
'YDL037C',
'YGL209W',
'YMR319C',
'YJR047C',
'YIL177C',
'YPR157W',
'YDR222W',
'YDR345C',
'YGR138C',
'YHR094C'
)

y = long_res_df %>%
  mutate(lysine = ifelse(str_detect('condition', 'plus'), 'plus', 'minus')) %>%
  dplyr::rename(ko = genotype) %>%
  dplyr::select(-condition)

x = cc_df %>% filter(
  id %in% glucose_upreg_genes,
  pval < .05 
) %>%
  group_by(ko, lysine) %>%
  arrange(desc(pval), .by_group = TRUE) %>%
  mutate(ko = toupper(ko)) %>%
  left_join(y)
```

Get genes affected by lysine

```{r}

```


# FINAL RESULTS

Eds1 has an activating effect on glycolysis genes, primarily in low glucose, high glucose, and galactose with lysine conditions.
Eds1 has a repressive effect on expression of glyoxylate cycle genes.--Combine glyoxylate and TCA networks
Eds1 has a mostly repressive effect on expression of TCA cycle genes.
It is not immediately clear what Eds1’s effect on fermentation pathway genes is.
Eds1 and Rgt1
Eds1 and Rgt1 directly regulate many common targets in the same direction.
According to single deletions--double deletions not yet processed because DE analysis pipeline needs to be moved to cluster.
Regulation of Eds1
There is no evidence in the single-deletion data for regulation of EDS1 by Rgt1, Mig1, or Mig2.
Westholm et al found that Mig1 and Mig2 coregulate EDS1--yet to be determined in our double deletions.
Lysine Biosynthesis Genes--Revist
Eds1 activates LYS4, ACO2, LYS9, LYS12, CTP1, LYS2, LYS1, LYS14 in low glucose -lysine.
LYS4 repressed in galactose +/-lysine, ACO2 repressed in galactose -lysine, LYS2 repressed in high glucose.
LYS4, LYS9, ACO2 are the only genes for which |LFC| > 1.
HXT Regulation network
Eds1 strongly and directly activates HXK1, where Rgt1 does not.
Eds1 directly represses MTH1 in galactose with lysine.
Rgt1 represses in galactose with lysine, galactose without lysine, high glucose late.
Mig1 represses in high glucose early, high glucose late.
Mig2 activates in high glucose early, represses in high glucose late.
Rgt1, Mig1, Mig2 all represses MTH1 according to Kuttykrishnan et al.
Eds1 indirectly represses STD1 in galactose with lysine and galactose without lysine.
Rgt1 represses STD1 in galactose with lysine, galactose without lysine, and low glucose.
Eds1 directly represses MIG1 in galactose no lysine, activates in glucose.
Rgt1 represses MIG1 in galactose +/- lysine.
Eds1 directly represses SKS1 in both galactose +/- lysine.
SKS1 is a repressor kinase of SNF3 (Yeast Genome), a low glucose sensor required for induction of certain hexose transporters and which triggers inhibition of Rgt1 (Ozcan et al, 1998).
Eds1 directly activates CDC19 in galactose +lysine.
CDC19 synthesizes pyruvate from phosphoenolpyruvate, the last step before TCA.
Rgt1 directly represses CDC19 in our -lysine conditions, but activates in galactose +lysine.
Phosphate Regulation
Eds1 causes highly significant repression of PHO11, PHO12, PHO89 (11-fold), SPL2 (32-fold) across all time points.
Also causes highly significant activation of PHO89 in galactose +lysine.
No evidence of direct binding.
Rgt1 also causes highly significant repression of these genes in glucose -lysine.
EDS1 deletion causes highly significant repression of PHO84 (11-fold) from T-1 to T90 glucose -lysine.
No evidence of direct binding.
Rgt1 also causes highly significant repression in glucose -lysine.
Mig1 activates in high glucose.
Eds1 directly activates TDA1 in our glucose -lysine conditions.
Protein of unknown function. Relocalizes from nuc to cytoplasm on DNA replication stress. RGT1, and to a lesser extent EDS1, do influence meiosis pathway genes.
Directly activated by RGT1.
Significant differential expression in Mig1 and Mig2 single deletions, but no evidence for direct binding.
Listed as a phosphate-responsive gene by Ljungdahl et al, 2012 in phosphate-starvation conditions.
Genes DE across all glucose -lysine timepoints in Eds1 deletion
List
PHO11: repressible acid phosphatase
HHF1: histone H4
PHO89: plasma membrane Na+/Pi cotransporter
YBR296C-A: putative protein with unknown function
HTA1: histone H2A
PHM6: protein of unknown function
KEI1: component of inositol phosphorylceramide synthase
YDR413C: dubious ORF
YDR433W: dubious ORF
SPS2: protein expressed during sporulation
YHL034W-A: dubious ORF
SPL2: downregulates low-affinity phosphate transport during phosphate limitation
PHO12: repressible acid phosphatase
YIR042C: putative protein with unknown function
MOG1: nuclear protein that interacts with GTP-Gsp1p, a protein involved in maintenance of nuclear organization
MGM101: protein with role in mitochondrial DNA recombinatorial repair
YLR076C: dubious ORF
SUR7: component of eisosomes, possibly anchoring them in plasma membrane
VPS75: binds histones and Rtt109, stimulating HAT activity
RPL36B: ribosomal 60s subunit protein
No significant GO terms were found for these genes
Transcription factors
=======
>>>>>>> 8fa902069a054d712991ad24f35a0a2007b11ed5
Eds1 directly activates Mig3 in high glucose -lysine.

```{r}
plotGeneCountCustom(dds, 'YER028C', 'MIG3')$combined
```

Eds1 directly represses MIG1 in galactose no lysine, activates in glucose.
(see plot above)

Rgt1 represses MIG1 in galactose +/- lysine.
(see plot above)

Eds1 directly represses MTH1 in galactose with lysine.
```{r}
plotGeneCountCustom(dds, 'YDR277C', 'MTH1')$combined
```

# NOTES

## diaxic shift
Following glucose exhaustion, cells switch to oxygenic respiration 
(a.k.a. "diauxic shift"), which allows catabolizing ethanol and the other 
carbon compounds via the TCA cycle and oxidative phosphorylation in the 
mitochondria.

MIG1 represses genes whose expression is shut off when glucose is present such 
as those enconding enzymes for sugars maltose, sucrose and galactose

When cells become limited for glucose, Mig1p is phosphorylated by the Snf1 kinase 
complex, composed of the Snf1 kinase catalytic subunit, the gamma subunit Snf4p, 
and a beta subunit encoded by SIP1, SIP2, or GAL83 (5). 

cite: https://services.bio.ifi.lmu.de:1047/diauxicGRN/


# NOTES

## diaxic shift
Following glucose exhaustion, cells switch to oxygenic respiration 
(a.k.a. "diauxic shift"), which allows catabolizing ethanol and the other 
carbon compounds via the TCA cycle and oxidative phosphorylation in the 
mitochondria.

MIG1 represses genes whose expression is shut off when glucose is present such 
as those enconding enzymes for sugars maltose, sucrose and galactose

When cells become limited for glucose, Mig1p is phosphorylated by the Snf1 kinase 
complex, composed of the Snf1 kinase catalytic subunit, the gamma subunit Snf4p, 
and a beta subunit encoded by SIP1, SIP2, or GAL83 (5). 

cite: https://services.bio.ifi.lmu.de:1047/diauxicGRN/

